---
title: "Meteolux"
author: "Bodart Céline, Al Sawaf Haya,Gonella Luca"
date: "2022-04-28"
output: html_document
---

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(GGally)
```

> Explore meteological data from the Findel station from 1947: Temperatures and rain precipitations

The data is available freely from the great website <https://data.public.lu>

## Fetch data

##### Download the CSV located at data.public.lu as [daily-meteorological-findel](https://data.public.lu/fr/datasets/daily-meteorological-parameters-luxembourg-findel-airport-wmo-id-06590/). Assign the name `findel_rain`

```{r}
findel_rain <- read_csv("data/06590-dailyparams-luxfindel-1947-ongoing.csv", locale=locale(encoding="latin1"))
findel_rain
```

```{block, opts.label = "tip"}
As usual, the values don't come as easy as it should. The supplied header is poorly encoded.

It is suggested to skip the first line and supply your own header based on the **Parameter description**:


- DXT (°C): DAILY MAXIMUM AIR TEMPERATURE AT 2 M – 23:46 UTC (D-1) TO 23:45 UTC (D)
- DNT (°C): DAILY MINIMUM AIR TEMPERATURE AT 2 M – 23:46 UTC (D-1) TO 23:45 UTC (D)
- DRR06_06 (mm): DAILY AMOUNT OF PRECIPITATION – 05:46 UTC (D) TO 05:45 UTC (D+1)

For the date format, use a proper `date` specification with `col_date()` to read this column properly
```

```{r}
findel_rain <- findel_rain %>% mutate(DATE= as.Date(findel_rain$DATE, format= "%d.%m.%Y" )) %>% rename("T°max (°C) " = "DXT (°C)", "T°min (°C)" = "DNT (°C)", "Rain (mm)" = "DRR06_06 (mm)")
```

##### Check the last lines of the tibble, which complete month have you?

Nous avons les valeurs de janvier 1947 à mars 2022

##### Check if we have all measurements per year.

```{r}
(2022-1947)*365+90+19
```

## Exploratory plots

##### Use `GGally::ggpairs` on the non-date columns to have a glance of the date

```{r}
GGally::ggpairs(findel_rain %>% select(-DATE))
```

##### Apart from the obvious relation between min/max temperatures within a day, why temperatures appear bimodal?

```{block, opts.label = "tip"}
If you need some insights to find out what is going on, you could display the max temperature distribution per season.

To do so:

- Extract the month from the `date` column, `ludbridate::month(date, label = TRUE, abbr = TRUE)` does a great job
- Collapse month per season, like `Winter = c("Dec", "Jan", "Feb")` and so forth
- Fill distribution per `season`

Conclude
```

```{r}
findel_rain <- findel_rain %>%mutate(MONTH=findel_rain$DATE %>% lubridate::month(label = TRUE, abbr = TRUE))
```

```{r}
Winter = c("Dec", "Jan", "Feb")
Summer = c("Jun", "Jul", "Aug")
Spring = c("Mar", "Apr", "May")
Fall = c("Sep", "Oct", "Nov")
```

##### Using the same new `season` column you created before, display the distribution per season of the daily temperature difference

-   Which season has the minimal temperature difference within a day?

```{r}

```

## Temperatures per decades

#### Break dates by decades

```{block, opts.label = "tip"}
`lubridate::year()` extracts a year from a date format, but then for creating a decade,

you need to think of working with the relevant module (`%%`).
```

Once you have the decades, you should expect to have as number of observations:

     decade     n
       <dbl> <int>
    1   1940  1096
    2   1950  3652
    3   1960  3653
    4   1970  3652
        [...]

```{r}
# Write your answer here
```

#### Plot the mean temperatures, minimum and maximum per decade

Now that you know how to create the decades, you can summarise by them both minimum and maximum temperatures.

```{block, opts.label = "tip"}
Keep only the complete decades, from 1950, to 2010.

Once summarised, you need to pivot the data to have them tidy for plotting
```

```{r}
# Write your answer here
```

##### How many Celsius degrees increase can you estimate over 60 years in Luxembourg

```{r}
# Write your answer here
```

## Rain precipitations

##### On which day, did the Findel station observe the top rain quantity?

```{r}
# Write your answer here
```

```{r}
# Write your answer here
```

##### Reproduce this [plot](https://twitter.com/Francois_Jobard/status/1440246462361726980) by François Jobart for the Findel station

![](https://pbs.twimg.com/media/E_zGK4nVUAsL7Du?format=png&name=900x900)

```{block, opts.label = "tip"}
We will follow the advice of [Luca Mathias](https://twitter.com/meteomathias/status/1440681768747626503?s=20) a luxembourgish meteorologist. For the summer, we keep June to August. It makes the filtering easier too.

Then compute per year, the sum of precipitations and the mean of min/max temperatures.

Display the mean temperature / rain for the period 1991-2020 as one big dot.

You can let down the dashed lines of "arid / semi-arid" descriptions.
```

```{r}
# Write your answer here
```

##### What can you conclude?

```{r}
# Write your answer here
```

##### Compute and plot the number of tropical nights/frozen days per year

```{block, opts.label = "tip"}
A tropical night is defined as a minimal temperature > 20°C

Frozen means that the maximum temperature remain below 0°C

You could add a linear trend with `geom_smooth()` to highlight the obvious
```

```{r}
# Write your answer here
```

##### Monthly temperatures from 2010-2021 to a 30 years referential period

We would like to mimic another plot by [François Jobart](https://twitter.com/Francois_Jobard/status/1479708565849923586?s=20&t=rqxYEY6G36TDe93vnWcnzQ)

![](https://pbs.twimg.com/media/FIj6-k2WUAMy4zL?format=png&name=medium)

The fact that 2021 alone has some standard deviation is weird. We will use 2010-2021 as test period

```{block, opts.label = "tip"}
- Start by selecting the years 1951-1980 and the last decade 2012-2021.
- Then categorised the `year` for being either the reference or the test decade
- Extract the day in year using ludridate::yday()`
- Compute both mean / std dev for the mean daily temperature per category and day in year
- Use `geom_ribbon()` for the std deviation
```

```{r}
# Write your answer here
```

We see that the plot shows a lot of daily variation.

However, the recent decade appears above the 30 years reference almost all year.

##### Display the distribution of reference/last decade and compute if the temperatures are significantly different. Justify your answer

```{r}
# Write your answer here
```

##### Smooth the temperature and standard deviation with a rolling mean

The details of how the smoothing was done is not supplied. But using a rolling mean we can obtain something close.

```{block, opts.label = "tip"}
Per reference/last decade, smooth use `slider::slide_mean` both the mean / sd temperatures.

You need to play with the `before` argument to find which period seems appropriate.

```

```{r}
# Write your answer here
```

## Distribution of rains

```{r}
# Write your answer here
```

##### Applied the same procedure for the 2 references but on the rain precipitations

-   Plot the distribution of the rain precipitations per reference (`ggbeeswarm::geom_quasirandom()` is great)
-   Test if the means of both the 30 years reference and last decade are different
-   Conclude

```{r}
# Write your answer here
```

## Acknowledgements

-   [Luca Mathias](https://twitter.com/meteomathias)
-   [François Jobart](https://twitter.com/Francois_Jobard)
